---
title: "Data visualization - part 3"
output: github_document
---

Load libraries that you will need:

```{r setup}
library(tidyverse)
library(viridis)
library(ggridges)
library(patchwork)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

## Load dataset
Load in a dataset that we will use often (copied from course webiste).
rnoaa has an online repository fo data

```{r}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728", "USC00519397", "USS0023B17S"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2017-01-01",
    date_max = "2017-12-31") %>%
  mutate(
    name = recode(
      id, 
      USW00094728 = "CentralPark_NY", 
      USC00519397 = "Waikiki_HA",
      USS0023B17S = "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10,
    month = lubridate::floor_date(date, unit = "month")) %>%
  select(name, id, everything())
```

## `group_by`

Note that you can use ungroup to undo the grouping
```{r}
weather_df %>% 
  group_by(name, month)
```

## counting things

count month / name observations
summarize by naming a variable and defining a function
```{r}
weather_df %>% 
  group_by(month) %>% 
  summarize(n_obs = n())

weather_df %>% 
  group_by(name) %>% 
  summarize(n_obs = n())

weather_df %>% 
  group_by(name, month) %>% 
  summarize(n_obs = n())
```

we can use`count()` as well if we are only counting things, does the same as above

```{r}
weather_df %>% 
  count(month, name = "n_obs")

weather_df %>% 
  count(name, month, name = "n_obs")
```

**NEVER** use base R's `table`
Problems: result is not a data frame, requires taking data out of a data frame

```{r, eval = FALSE}
weather_df %>% 
  pull(month) %>% 
  table()
```

Other helpful counters

n_distinct tells you how many unique dates there are in each category 

```{r}
weather_df %>% 
  group_by(month) %>% 
  summarize(
    n_obs = n(),
    n_days = n_distinct(date))
```

## A digression on 2x2 tables (exposure with 2 levels, response with 2 levels)

TRUE ~ "" accounts for what to do with values that are missed by the aforementioned conditions

```{r}
weather_df %>% 
  filter(name != "Waikiki_HA") %>% 
  mutate(
    cold = case_when(
      tmax < 5 ~ "cold",
      tmax >= 5 ~ "not_cold",
      TRUE  ~ "")
    ) %>% 
      group_by(name, cold) %>% 
      summarize(count = n())
```

```{r}
weather_df %>% 
  filter(name != "Waikiki_HA") %>% 
  mutate(
    cold = case_when(
      tmax < 5 ~ "cold",
      tmax >= 5 ~ "not_cold",
      TRUE  ~ "")
    ) %>% 
janitor::tabyl(name, cold)
```


## General summaries

Take the mean tmax
na.rm eliminates "NA" values

```{r}
weather_df %>% 
  group_by(month) %>% 
  summarize(
    mean_tmax = mean(tmax, na.rm = TRUE),
    mean_prcp = mean(prcp, na.rm = TRUE),
    median_tmin = median(tmin, na.rm = TRUE)
  )

weather_df %>% 
  group_by(name, month) %>% 
  summarize(
    mean_tmax = mean(tmax, na.rm = TRUE),
    mean_prcp = mean(prcp, na.rm = TRUE),
    median_tmin = median(tmin, na.rm = TRUE)
  )
```

This is a dataframe!

```{r}
weather_df %>% 
  group_by(name, month) %>% 
  summarize(
    mean_tmax = mean(tmax, na.rm = TRUE),
    mean_prcp = mean(prcp, na.rm = TRUE),
    median_tmin = median(tmin, na.rm = TRUE)
  ) %>% 
  ggplot(aes(x = month, y = mean_tmax, color = name)) +
  geom_point() +
  geom_line()
```

Alternatively, if you want to see each individual data point and lines connecting them, you can use:

```{r}
weather_df %>% 
  ggplot(aes(x = date, y = tmax, color = name)) +
  geom_point() +
  geom_line()
```

