---
title: "Data visualization - part 2"
output: github_document
---

Load libraries that you will need:

```{r setup}
library(tidyverse)
library(viridis)
library(ggridges)
library(patchwork)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)
```

## Setting options
Setting options that can be included at the beginning of any R Markdown document
We won't load now because it will affect later examples in this document

library(tidyverse)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d

## Load dataset
Load in a dataset that we will use often (copied from course webiste).
rnoaa has an online repository fo data

```{r}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728", "USC00519397", "USS0023B17S"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2017-01-01",
    date_max = "2017-12-31") %>%
  mutate(
    name = recode(
      id, 
      USW00094728 = "CentralPark_NY", 
      USC00519397 = "Waikiki_HA",
      USS0023B17S = "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>%
  select(name, id, everything())
```

Create a scatterplot and modify labels, colors, scales, etc.

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax)) +
  geom_point(aes(color = name), alpha = .3) +
  labs(
    title = "Temperature at three stations",
    x = "Minimum daily temp (C)",
    y = "Maximum daily temp (C)",
    caption = "Data from rnoaa package with three stations") +
  scale_x_continuous(
    breaks = c(-15, 0, 15),
    labels = c("-15 C", "0 C", "15 C")
  ) + 
  scale_y_continuous(
    trans = "sqrt",
    position = "right"
  ) +
  scale_color_hue(
    name = "Location",
    h = c(100,300))

weather_df %>% 
  ggplot(aes(x = tmin, y = tmax)) +
  geom_point(aes(color = name), alpha = .3) +
  labs(
    title = "Temperature at three stations",
    x = "Minimum daily temp (C)",
    y = "Maximum daily temp (C)",
    caption = "Data from rnoaa package with three stations") +
  scale_x_continuous(
    breaks = c(-15, 0, 15),
    labels = c("-15 C", "0 C", "15 C")
  ) + 
  scale_color_hue(
    name = "Location",
    h = c(100,300)) +
  scale_color_viridis_d()
```

Viridis color palette is very distinct visually (load it in using package above)
"_d" specifies that it is a discrete variable

## Themes

Plot elements that are not directly tied to data or geometry, just might make the plot look nicer

Note that order matters for some of these functions. If you update the theme to bw before you move the legend, the legend won't move.

Themes: _bw, _minimal, _classic (many more available on the internet)

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax)) +
  geom_point(aes(color = name), alpha = .3) +
  labs(
    title = "Temperature at three stations",
    x = "Minimum daily temp (C)",
    y = "Maximum daily temp (C)",
    caption = "Data from rnoaa package with three stations") +
  scale_color_viridis_d(
    name = "Location") +
  theme_bw() +
  theme(legend.position = "bottom")

weather_df %>% 
  ggplot(aes(x = tmin, y = tmax)) +
  geom_point(aes(color = name), alpha = .3) +
  labs(
    title = "Temperature at three stations",
    x = "Minimum daily temp (C)",
    y = "Maximum daily temp (C)",
    caption = "Data from rnoaa package with three stations") +
  scale_color_viridis_d(
    name = "Location") +
  theme_classic() +
  theme(legend.position = "bottom")

weather_df %>% 
  ggplot(aes(x = tmin, y = tmax)) +
  geom_point(aes(color = name), alpha = .3) +
  labs(
    title = "Temperature at three stations",
    x = "Minimum daily temp (C)",
    y = "Maximum daily temp (C)",
    caption = "Data from rnoaa package with three stations") +
  scale_color_viridis_d(
    name = "Location") +
  theme_minimal() +
  theme(legend.position = "bottom")
```

## `data` in geoms

```{r}
weather_df %>% 
  ggplot(aes(x = date, y = tmax, color = name)) +
  geom_point()

central_park = 
  weather_df %>% 
  filter(name == "CentralPark_NY")

waikiki =
  weather_df %>% 
  filter(name == "Waikiki_HA")

waikiki %>% 
  ggplot(aes(x = date, y = tmax, color = name)) +
  geom_point() + 
  geom_line()

waikiki %>% 
  ggplot(aes(x = date, y = tmax, color = name)) +
  geom_point() + 
  geom_line(data = central_park)
```

## `patchwork`
Requires the patchwork library
Method for displaying multiple separate plots at once

```{r}
ggp_tmax_tmin = 
  weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = .3) +
  theme(legend.position = "none")

ggp_prcp_dens = 
  weather_df %>% 
  filter(prcp > 0) %>% 
  ggplot(aes(x = prcp, fill = name)) +
  geom_density(alpha = .3) +
  theme(legend.position = "none")

ggp_tmax_date = 
  weather_df %>% 
  ggplot(aes(x = date, y = tmax, color = name)) +
  geom_point() +
  geom_smooth() +
  theme(legend.position = "bottom")

(ggp_tmax_tmin + ggp_prcp_dens)/ ggp_tmax_date
```

## data manipulation

Quick example on factor variables

```{r}
weather_df %>% 
  ggplot(aes(x = name, y = tmax)) +
  geom_boxplot()
```

Reorder name variable by tmax:

```{r}
weather_df %>% 
  mutate(
    name = fct_reorder(name, tmax)
  ) %>% 
  ggplot(aes(x = name, y = tmax)) +
  geom_boxplot()
```

What about tmax and tmin...
This requires transformation of the dataset using pivot_longer
This allows for temperature on the x-axis and fill with another variable
Remember that facet_grid allows for creation of separate plots stratified by a categorical variable
```{r}
weather_df %>% 
  pivot_longer(
    tmax:tmin,
    names_to = "obs",
    values_to = "temperature"
  ) %>% 
  ggplot(aes(x = temperature, fill = obs)) +
  geom_density(alpha = .3) +
  facet_grid(.~name)
```

This code chunk allows for creation of a box plot showing distribution of bdi at each visit
Therefore, you need a visit variable and a bdi variable
```{r}
pulse_df =
  haven::read_sas("data/public_pulse_data.sas7bdat") %>% 
  janitor::clean_names() %>% 
  pivot_longer(
    bdi_score_bl:bdi_score_12m,
    names_to = "visit",
    values_to = "bdi",
    names_prefix = "bdi_score_"
  ) %>% 
  mutate(visit = recode(visit, "bl" = "00m"))

pulse_df %>% 
  ggplot(aes(x = visit, y = bdi)) +
  geom_boxplot()
```

This creates lines for each individaul in the data set over time.
```{r}
pulse_df %>% 
  ggplot(aes(x = visit, y = bdi)) +
  geom_point(size = .2) + 
  geom_line(aes(group = id))
```

One more example using pups data set, studying fetal alcohol syndrome
```{r}
pup_data = 
  read_csv("./data/FAS_pups.csv", col_types = "ciiiii") %>%
  janitor::clean_names() %>%
  mutate(sex = recode(sex, `1` = "male", `2` = "female")) 

litter_data = 
  read_csv("./data/FAS_litters.csv", col_types = "ccddiiii") %>%
  janitor::clean_names() %>%
  separate(group, into = c("dose", "day_of_tx"), sep = 3)

fas_data = left_join(pup_data, litter_data, by = "litter_number") 

fas_data %>% 
  select(sex, dose, day_of_tx, pd_ears:pd_walk) %>% 
  pivot_longer(
    pd_ears:pd_walk,
    names_to = "outcome", 
    values_to = "pn_day") %>% 
  drop_na() %>% 
  mutate(outcome = forcats::fct_reorder(outcome, pn_day, median)) %>% 
  ggplot(aes(x = dose, y = pn_day)) + 
  geom_violin() + 
  facet_grid(day_of_tx ~ outcome)
```

