---
title: "Data visualization - ggplot 1"
output: github_document
---

Load libraries that you will need:

```{r setup}
library(tidyverse)
library(ggridges)
```

Load in a dataset that we will use often (copied from course webiste).
rnoaa has an online repository fo data

```{r}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728", "USC00519397", "USS0023B17S"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2017-01-01",
    date_max = "2017-12-31") %>%
  mutate(
    name = recode(
      id, 
      USW00094728 = "CentralPark_NY", 
      USC00519397 = "Waikiki_HA",
      USS0023B17S = "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>%
  select(name, id, everything())
```

## Scatterplot

tmax vs. tmin
create a ggplot object

```{r}
ggp_tmax_tmin =
  weather_df %>% 
  ggplot(aes(x = tmin, y = tmax)) +
  geom_point()

ggp_tmax_tmin
```

## Let's fancy it up

Add ... color? lines? other stuff?
geom_smooth(se = FALSE) adds a line for each color
geom_point with alpha blending changes the transparency of the dots, makes it easier to see the lines

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = .3) +
  geom_smooth(se = FALSE)
  
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax)) + 
  geom_point(aes(color = name), alpha = .5) +
  geom_smooth(se = FALSE)

weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = .3) +
  geom_smooth(se = FALSE) + 
  facet_grid(. ~ name)

```

