---
title: "Data visualization - ggplot 1"
output: github_document
---

Load libraries that you will need:

```{r setup}
library(tidyverse)
library(ggridges)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)
```

Load in a dataset that we will use often (copied from course webiste).
rnoaa has an online repository fo data

```{r}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728", "USC00519397", "USS0023B17S"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2017-01-01",
    date_max = "2017-12-31") %>%
  mutate(
    name = recode(
      id, 
      USW00094728 = "CentralPark_NY", 
      USC00519397 = "Waikiki_HA",
      USS0023B17S = "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>%
  select(name, id, everything())
```

## Scatterplot

tmax vs. tmin
create a ggplot object

```{r}
ggp_tmax_tmin =
  weather_df %>% 
  ggplot(aes(x = tmin, y = tmax)) +
  geom_point()

ggp_tmax_tmin
```

## Let's fancy it up

Add ... color? lines? other stuff?
geom_smooth(se = FALSE) adds a line for each color
geom_point with alpha blending changes the transparency of the dots, makes it easier to see the lines

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = .3) +
  geom_smooth(se = FALSE)
  
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax)) + 
  geom_point(aes(color = name), alpha = .5) +
  geom_smooth(se = FALSE)

weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = .3) +
  geom_smooth(se = FALSE) + 
  facet_grid(. ~ name)
```

Let's make one more scatterplot.

```{r}
weather_df %>% 
  ggplot(aes(x = date, y = tmax)) +
  geom_point() +
  facet_grid(. ~name)

weather_df %>% 
  ggplot(aes(x = date, y = tmax)) +
  geom_point(alpha = .3) +
  facet_grid(. ~name) +
  geom_smooth(se = FALSE)

weather_df %>% 
  ggplot(aes(x = date, y = tmax, size = prcp)) +
  geom_point(alpha = .3) +
  facet_grid(. ~name) +
  geom_smooth(se = FALSE)
```

se = FALSE turns off error bars

## Use data manipulation as part of this

```{r}
weather_df %>% 
  filter(name == "CentralPark_NY") %>% 
  mutate(
    tmax = tmax * (9/5) + 32,
    tmin = tmin * (9/5) + 32
  ) %>% 
  ggplot(aes(x = date, y = tmax)) +
  geom_point()
```

## Stacking geoms

Which geoms do you want?

```{r}
weather_df %>% 
  ggplot(aes(x = date, y = tmax, color = name)) +
  geom_smooth()
```

Geom_hex (requires "hexbin" package)
```{r}
weather_df %>% 
  ggplot(aes(x = tmax, y = tmin)) + 
  geom_hex()
```

Histograms
```{r}
weather_df %>% 
  ggplot(aes(x = tmax)) + 
  geom_histogram()

weather_df %>% 
  ggplot(aes(x = tmax, fill = name)) + 
  geom_histogram()

weather_df %>% 
  ggplot(aes(x = tmax, fill = name)) + 
  geom_histogram() +
  facet_grid(. ~name)
```

Let's try some other plots...

geom_density gives a smoothed out line

```{r}
weather_df %>% 
  ggplot(aes(x = tmax)) + 
  geom_density()

weather_df %>% 
  ggplot(aes(x = tmax, fill = name)) + 
  geom_density()

weather_df %>% 
  ggplot(aes(x = tmax, fill = name)) + 
  geom_density(alpha = .3)
```

Still with 'tmax' and 'name'
box plots
```{r}
weather_df %>% 
  ggplot(aes(x = name, y = tmax)) +
  geom_boxplot()
```

Some people like violin plots????

```{r}
weather_df %>% 
  ggplot(aes(x = name, y = tmax)) +
  geom_violin()
```

What about ridges...

```{r}
weather_df %>% 
  ggplot(aes(x = tmax, y = name)) +
  geom_density_ridges()

weather_df %>% 
  ggplot(aes(x = tmax, y = name)) +
  geom_density_ridges(alpha = .8, scale = .8)
```

## Embedding plots

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = .3)
```

```{r, fig.width = 7, fig.height = 5}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = .3)
```

## Learning Assessments

```{r}
weather_df %>% 
  filter(name == "CentralPark_NY") %>% 
  mutate(
    tmax = tmax * (9/5) + 32,
    tmin = tmin * (9/5) + 32
  ) %>% 
  ggplot(aes(x = tmin, y = tmax)) +
  geom_point(alpha = .5) +
  geom_smooth(method = "lm", se = FALSE)
```

```{r}
ggplot(weather_df) + geom_point(aes(x = tmax, y = tmin), color = "blue")
ggplot(weather_df) + geom_point(aes(x = tmax, y = tmin, color = "blue"))
```

Make plots that compare precipitation across locations. Try a histogram, a density plot, a boxplot, a violin plot, and a ridgeplot; use aesthetic mappings to make your figure readable.

```{r}
weather_df %>% 
  ggplot(aes(x = prcp, fill = name)) + 
  geom_histogram()

weather_df %>% 
  ggplot(aes(x = prcp, fill = name)) + 
  geom_histogram() +
  facet_grid(. ~name)

weather_df %>% 
  ggplot(aes(x = prcp, fill = name)) + 
  geom_density()

weather_df %>% 
  ggplot(aes(x = prcp)) + 
  geom_density(aes(fill = name), alpha = .5)

weather_df %>% 
  ggplot(aes(x = name, y = prcp)) +
  geom_boxplot()

weather_df %>% 
  ggplot(aes(x = name, y = prcp)) +
  geom_violin()

weather_df %>% 
  ggplot(aes(x = prcp, y = name)) +
  geom_density_ridges(scale = .85)

weather_df %>% 
  filter(prcp > 0) %>% 
  ggplot(aes(x = prcp, y = name)) + 
  geom_density_ridges(scale = .85)

```

